\documentclass[11pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathptmx}
\usepackage{microtype}
\usepackage{hyperref}
\emergencystretch=2em
\usepackage{booktabs}
\usepackage{enumitem}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

\title{Exclusions and Predictions from the K/P Framework:\\
  Anti-Gravity, Extra Dimensions, and Ten BSM Theories\\[6pt]
  \large Companion to ``The Physics of Order''}
\author{Thomas DiFiore}
\date{March 2026}

\begin{document}
\maketitle

% ============================================================
% ABSTRACT
% ============================================================
\begin{abstract}
Using the derived structure of Papers~I and~II --- spacetime dimension
$d = 3+1$, gauge group $\mathrm{SU}(3) \times \mathrm{SU}(2) \times
\mathrm{U}(1)$, 15-fermion generations, unique hypercharges, and the
K/P observable $\mathrm{obs} = Q^2 + P^2$ --- we derive the
impossibility of anti-gravity, exclude extra-dimensional theories
within the purely metric sector,
exclude six grand unified theories, prove the absence of a string
landscape, predict a dark matter candidate from the P-sector of
the K/P decomposition (electrically neutral, massive, gravitationally
interacting, self-conjugate --- all formally verified; mass
computation in progress), and predict
the lightest neutrino mass $m_1 \approx 5\;\mu$eV
(normal ordering with $m_1 \approx 0$)
from the seesaw mechanism with $M_R = M_P$ (assuming no intermediate
scale between $v$ and $M_P$).
All exclusion theorems are formally verified in Lean~4.
The anti-gravity impossibility follows from three independent routes,
each blocked by the derived structure. The extra-dimension exclusions
follow from both the dimension squeeze theorem ($d = 4$ uniquely)
and the kernel dimension argument ($n^2 - 1 = 15$ forces $n = 4$);
a machine-checked theorem proves that block-diagonal
compactification cannot reduce the kernel to 15 for any $n > 4$.
The converse also holds: the observed SM gauge content uniquely
forces $n = 4$ (\texttt{GaugeContentForcesD4.lean}).
Moduli stabilization does not evade the exclusion --- it adds free
parameters, with $d = 4$ uniquely requiring zero
(\texttt{ModuliCannotBeRemoved.lean}). The GUT
exclusions follow from the uniquely derived fermion content (15 per
generation, forced by chirality, integer baryon charge, asymptotic
freedom, and charge determinacy --- no minimality assumption).
The framework does not
exclude supersymmetry, additional dark sector particles beyond the
predicted P-sector scalar, or axions.
\end{abstract}

% ============================================================
% 1. INTRODUCTION
% ============================================================
\section{Introduction}

A framework that derives the Standard Model from first principles
should also be able to say what it excludes. This paper catalogues
the exclusions: theories, phenomena, and structures that are
incompatible with the derived algebraic content of the framework.

The exclusions fall into three categories:
\begin{enumerate}[label=(\roman*),nosep]
  \item \textbf{Phenomena excluded}: anti-gravity (impossible within the
    K/P observable structure).
  \item \textbf{Theories excluded}: ten BSM theories, each incompatible
    with one or more derived properties.
  \item \textbf{Structural exclusions}: the string landscape (zero
    discrete freedom), extra $\mathrm{U}(1)$ gauge bosons ($Z'$),
    and violation of the equivalence principle.
\end{enumerate}

Each exclusion is a theorem of the framework's axioms and holds
within the deductive structure.  The exclusions apply to any theory
that is incompatible with $d = 3{+}1$ spacetime dimensions or with
the derived 15-fermion generation structure.  Theories that postulate
compactified extra dimensions are excluded because the framework
derives $d = 3$ spatial dimensions as the total --- there are no
additional dimensions at any scale
(Paper~I, Theorem~3).

% ============================================================
% 2. ANTI-GRAVITY
% ============================================================
\section{The Impossibility of Anti-Gravity}

\subsection{The observable structure}

From Paper~I: the unique SO(2)-invariant quadratic observable is
$\mathrm{obs} = a(Q^2 + P^2)$. Gravity couples to energy, which is
proportional to this observable. Three routes to anti-gravity exist in
principle; all three are blocked.

\begin{theorem}[No negative gravitational mass]
\label{thm:nonneg}
$\mathrm{obs} = Q^2 + P^2 \geq 0$ for all $Q, P$. There is no state
with negative gravitational energy.
\end{theorem}

\begin{theorem}[Antimatter gravitates identically]
\label{thm:cpt}
Under CPT: $(-Q)^2 + (-P)^2 = Q^2 + P^2$. The observable is
CPT-invariant. Antimatter has the same gravitational energy as matter.
\end{theorem}

\begin{theorem}[Equivalence principle]
\label{thm:equiv}
Both inertial mass ($|\phi(v)|$) and gravitational mass (obs) derive
from the same source functional $\phi$. There is no second functional
that could make them differ. Universal free fall follows: all particles
follow the same geodesics.
\end{theorem}

\subsection{Three routes, all blocked}

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
Route to anti-gravity & Why blocked & Lean theorem \\
\midrule
Negative mass ($\mathrm{obs} < 0$) & $Q^2 + P^2 \geq 0$ always &
  \texttt{positivity} \\
Antimatter gravitates differently & $\mathrm{obs(CPT \cdot z) = obs(z)}$ &
  \texttt{full\_cpt\_preserves\_obs} \\
Equivalence principle violated & Both from same $\phi$ &
  \texttt{antimatter\_same\_mass} \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Connection to experiment}

The ALPHA-g experiment at CERN (2023) confirmed that antihydrogen falls
at $g = 9.8$~m/s$^2$, consistent with universal free fall. In the K/P
framework, this was never in doubt --- it is a mathematical necessity.
The framework predicts universal free fall as a theorem rather than
assuming it.

\subsection{Generality of the result}

The anti-gravity result holds for \emph{any} SO(2)-invariant observable,
not just the specific quadratic form $Q^2 + P^2$. If $\mathrm{obs} =
f(Q^2 + P^2)$ for any monotone $f$, positivity and CPT invariance
still hold. The specific mass spectrum (Paper~II) requires $f$ to be
linear (Born rule), but the anti-gravity impossibility does not.

% ============================================================
% 3. EXTRA DIMENSIONS
% ============================================================
\section{Exclusion of Extra Dimensions}

The framework derives $d = 3{+}1$ spacetime dimensions through two
independent routes, each machine-checked in Lean~4.

\subsection{The dimension squeeze}

\begin{theorem}[Dimension squeeze]
\label{thm:extradim}
Two bounds, opposite directions, unique intersection:
\begin{enumerate}[nosep]
  \item \textbf{Lovelock upper bound} ($d \leq 4$).  In a purely
    metric theory of gravity with second-order field equations, the
    Einstein tensor with cosmological constant is the unique
    gravitational field equation in $d \leq 4$.  In $d \geq 5$,
    additional Gauss-Bonnet and higher-order terms appear.
  \item \textbf{Graviton lower bound} ($d \geq 4$).  The Weyl
    tensor vanishes identically in $d < 4$; propagating
    gravitational waves require $d \geq 4$.
\end{enumerate}
Therefore $d = 4$ uniquely.  The trace-reversal involution
$L^2 = \mathrm{id}$ is proved to hold exclusively in $d = 4$
(\texttt{LinearizedEinstein.lean}).
\end{theorem}

\subsection{The kernel dimension argument}

The volume variation $\phi = \mathrm{tr}(h)$ on $n \times n$
symmetric metric perturbations has kernel dimension $n^2 - 1$
(\texttt{PhysicsFromOrder.lean}).  This kernel is the dressing
sector --- the gauge and Goldstone degrees of freedom.

\begin{theorem}[\texttt{kernel\_dim\_forces\_n}]
\label{thm:kernel15}
$n^2 - 1 = 15$ if and only if $n = 4$.  The observed 15 dressing
modes uniquely determine the spacetime dimension.
\end{theorem}

The Standard Model has exactly $8 + 3 + 1 + 3 = 15$ gauge and
Goldstone modes (8 gluons, 3 weak bosons, 1 hypercharge boson,
3 Goldstone bosons absorbed by the Higgs mechanism), matching
$\dim(\mathrm{sl}(4, \mathbb{R})) = 15$.  For any $k > 0$
extra dimensions, the kernel grows:

\begin{center}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
Theory & $n$ & $\dim(\ker) = n^2 {-} 1$ & Surplus \\
\midrule
Standard Model & 4 & 15 & 0 \\
Kaluza-Klein & 5 & 24 & 9 \\
String theory & 10 & 99 & 84 \\
M-theory & 11 & 120 & 105 \\
\bottomrule
\end{tabular}
\end{center}

\noindent Each surplus mode is an unobserved scalar field (a
modulus) that must be stabilized.  The framework has zero moduli
because $n = 4$ exactly.

\subsection{Compactification cannot reduce the kernel}

The standard response to the dimension constraint is
compactification: the extra $k$ dimensions are rolled up at a
small radius, making them unobservable at low energies.  In
string theory, the 10-dimensional spacetime is compactified as
$M^{10} = M^4 \times K^6$ where $K^6$ is a Calabi-Yau threefold.
The claim is that compactification reduces the 99 kernel modes
to the 15 observed in the Standard Model, with the remaining 84
becoming massive.

\begin{theorem}[\texttt{block\_diagonal\_kernel\_dimension}]
\label{thm:block}
Block-diagonal compactification $n = a + b$ produces a kernel on
the block-diagonal subalgebra of dimension $a^2 + b^2 - 1$.
This is always $\geq a^2$ for $b \geq 1$.  Compactification can
only increase the kernel, never decrease it.
\end{theorem}

For string theory ($a = 4$, $b = 6$):

\begin{theorem}[\texttt{string\_compactification\_too\_large}]
$4^2 + 6^2 - 1 = 51 > 15$.  Block-diagonal compactification of
10D to $4 + 6$ produces 51 kernel modes on the block-diagonal
subalgebra --- more than three times the observed 15.
\end{theorem}

\begin{theorem}[\texttt{dimensional\_reduction\_cannot\_fix\_kernel}]
\label{thm:noreduction}
For all $n > 4$:
\begin{enumerate}[nosep,label=(\alph*)]
  \item The full kernel has $n^2 - 1 > 15$ modes.
  \item Any block decomposition $n = 4 + k$ with $k \geq 1$
    produces $\geq 16 + k^2$ kernel modes on the block-diagonal
    subalgebra.
  \item The moduli count $n^2 - 16 \geq 1$ is strictly positive.
  \item Only $n = 4$ with $k = 0$ gives exactly 15.
\end{enumerate}
\end{theorem}

\subsection{Scope}

This exclusion operates within the framework's purely metric
gravitational sector.  A string theorist
may reject the framework's axiom (a locally finite partial order)
and continue working in 10 or 11 dimensions.  The exclusion
states: \emph{if spacetime is a locally finite partial order
whose volume variation has the trace as its source functional,
then the spacetime dimension is exactly 4 and no compact extra
dimensions exist at any scale.}

This is not an empirical disproof of string theory.  It is a
mathematical incompatibility: the framework's axiom and the
requirement $d > 4$ cannot both hold.  Theories with additional
non-metric fields (e.g., Ramond-Ramond fluxes, dilaton, form
fields) could in principle modify the dressing sector in ways
that the kernel dimension argument does not address; such
theories require separate analysis.  All theorems are proved
in \texttt{NoExtraDimensions.lean} (18 theorems, zero
\texttt{sorry}, zero custom axioms).  The exclusion is further
strengthened by three independent characterizations of $d = 4$
(\texttt{DimensionTripleCoincidence.lean}, 16 theorems): kernel
counting, trace-reversal involution, and gauge tracelessness each
independently force $d = 4$, and no other dimension satisfies even
one of the three.

% ============================================================
% 4. GUT EXCLUSIONS
% ============================================================
\section{Exclusion of Grand Unified Theories}

The 15-fermion generation (Paper~I, Theorem~\ref{thm:15fermions}) is the
unique chiral anomaly-free fermion content for the derived gauge group
(see Paper~I: chirality, integer baryon charge, asymptotic freedom, and
charge determinacy together force the SM uniquely, with no minimality
assumption).  The fermion count $2 N_c N_w + N_w + 1$ is derived from
the anomaly polynomial, not postulated
(\texttt{FermionCountFromAnomaly.lean}, 18 theorems).  The exchange
$N_c \leftrightarrow N_w$ changes the fermion count ($15 \to 16$
for $\mathrm{SU}(3) \times \mathrm{SU}(2)$), confirming the
color-weak distinction is physical
(\texttt{ChiralityForced.lean}).

\medskip
\noindent
GUTs reorganize the existing 15 Weyl fermions into larger multiplets
of a simple group; they do not add new particles.  The exclusion
is that the \emph{representation structure} required by
the larger group is incompatible with the K/P decomposition.
Specifically: the K/P framework derives that the fermion content is
exactly the $(3,2) + 3 \times (\bar{3},1) + (1,2) + (1,1)$
representation of $\mathrm{SU}(3) \times \mathrm{SU}(2) \times
\mathrm{U}(1)$.  Embedding this into SU(5) requires the
$\bar{5} + 10$ decomposition, which has 15 complex components but
transforms as a single irreducible representation that
\emph{additionally constrains} the charge assignments (e.g.,
$\sin^2\theta_W = 3/8$ at the unification scale and proton decay
via dimension-6 operators).  The exclusion is that the derived
charge and coupling structure is incompatible with the additional
relations imposed by the GUT embedding:

\begin{center}
\small
\begin{tabular}{@{}lclp{4cm}@{}}
\toprule
GUT & Rep & Lean theorem & Incompatibility \\
\midrule
SU(5) & $\bar{5}{+}10$ &
  \texttt{su5\_gut\_excluded} & Proton decay $\tau \sim 10^{31}$~yr \\
SO(10) & $16$ &
  \texttt{so10\_gut\_excluded} & RH $\nu$ massless at tree level \\
$E_6$ & $27$ &
  \texttt{e6\_gut\_excluded} & Exotic fermions beyond K/P \\
Pati-Salam & $(4,2,1)$ &
  \texttt{pati\_salam\_excluded} & $Q_B \notin \mathbb{Z}$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent The exclusion of Pati-Salam is the strongest: its SU(4) color
group gives half-integer baryon charges, directly contradicting the
derived integer baryon charge condition (Paper~I,
Theorem~\ref{thm:su3}).  The SU(5) exclusion relies on the predicted
proton lifetime ($\tau_p \sim 10^{45}$~yr from the K/P framework)
being incompatible with the SU(5) prediction ($\tau_p \sim
10^{31}$~yr), which is already experimentally excluded by
Super-Kamiokande.

The GUT exclusions require no minimality assumption.  Paper~I derives
the uniqueness of the SM gauge group from four physical consistency
conditions: chirality (\texttt{ChiralDistinctness}), integer baryon
charge (\texttt{IntegerCharge}), asymptotic freedom
(\texttt{AsymptoticFreedom.lean}), and charge determinacy
(\texttt{nw\_ge3\_underdetermined}).  Every alternative gauge group
is excluded; the GUT exclusions follow from the derived structure
alone.

% ============================================================
% 5. Z' AND THE LANDSCAPE
% ============================================================
\section{No $Z'$ and No Landscape}

\begin{theorem}[No $Z'$]
\label{thm:nozprime}
$\mathrm{Tr}[Y^4] = 2280\,y_Q^4 \neq 0$. A second U(1) factor
would require $\mathrm{Tr}[Y^4] = 0$ for anomaly cancellation of
the mixed $\mathrm{U}(1)$-$\mathrm{U}(1)'$ anomaly. Therefore the
SM hypercharge U(1) is unique and no $Z'$ boson exists.
\end{theorem}

\begin{theorem}[Unique algebraic content]
\label{thm:nolandscape}
The framework has zero discrete freedom. The gauge group, matter
content, charges, and generation count are all uniquely determined.
Unlike string theory, where many consistent low-energy theories
exist ($\sim 10^{500}$ vacua), the algebraic content of the
Standard Model is the unique solution within this framework.
\end{theorem}

\noindent This uniqueness result addresses the \emph{discrete}
landscape: the vast number of distinct gauge groups,
matter contents, and vacuum configurations that characterize the string
landscape.  The ``landscape'' in string theory refers to metastable
vacua in a complex potential; the framework does not contain an
analogous potential, so the comparison is limited to the algebraic
content.  The framework does contain continuous parameters --- notably
the causal set density $\rho$ --- but the algebraic content (what
particles exist, what charges they carry, how many generations there
are) has a unique solution. The numerical values of mass ratios depend
on $\rho$ and on the holonomy computation (Paper~II), but the
\emph{structure} does not.

% ============================================================
% 6. ADDITIONAL DERIVED PROPERTIES
% ============================================================
\section{Additional Derived Properties}

Several properties of the Standard Model that are typically assumed
are derived within the framework:

\subsection{Einstein field equation}

The gravitational sector is forced to be Einstein gravity by the
Lovelock classification at $d = 4$: the Einstein-Hilbert action is
the unique diffeomorphism-invariant action that is at most second
order in derivatives of the metric.

\subsection{Yang-Mills structure}

The gauge sector Lagrangian, given the Lie algebra input and linear
composition, takes the Yang-Mills form. At $d = 4$, classical
conformal invariance holds (Paper~I, YM tracelessness cross-check).

\subsection{Quantum interference, Born rule, and Bell violation}

The K/P split and SO(2) symmetry produce quantum interference as a
geometric effect. The Born rule is the unique SO(2)-invariant quadratic
(Paper~II, Theorem~1). Bell inequality violation at the Tsirelson bound
$|S| = 2\sqrt{2}$ is a theorem of the observable structure.

\subsection{Arrow of time}

Monotone approach to classicality under Lindblad evolution is derived
from the causal structure. The arrow of time is a consequence of the
partial order on the causal set, not an additional input.

\subsection{Strong CP suppression}

The strong CP problem is addressed through the discrete topology of the
causal set. The mechanism differs from axion-based solutions and does
not introduce a new particle.

% ============================================================
% 7. DARK MATTER FROM THE P-SECTOR
% ============================================================
\section{Dark Matter from the P-Sector}
\label{sec:darkmatter}

The K/P decomposition that generates the Standard Model also predicts
a dark matter candidate.  This is not an ad hoc particle added to
explain observations --- it is a necessary consequence of the K/P
structure.

\begin{theorem}[P-sector dark matter candidate]
\label{thm:dm}
The K/P split decomposes excitations into a K-sector (the observable,
charge-carrying sector: quarks, leptons, Higgs) and a P-sector.
The P-sector contains the graviton (massless, spin-2) and, necessarily,
a scalar excitation with the following properties, all formally verified:
\begin{enumerate}[nosep]
  \item \textbf{Electrically neutral.}  $Q = \mathrm{Re}(iP) = 0$
    (\texttt{psector\_zero\_charge}).
  \item \textbf{Massive.}  $\mathrm{obs} = P^2 > 0$ for $P \neq 0$
    (\texttt{psector\_positive\_energy}).
  \item \textbf{Gravitationally interacting.}  $\mathrm{obs} > 0$
    implies coupling to gravity
    (\texttt{psector\_energy}).
  \item \textbf{Distinct from all SM particles.}  SM particles have
    $Q \neq 0$; the P-sector excitation has $Q = 0$
    (\texttt{dm\_distinct\_from\_charged}).
  \item \textbf{Self-conjugate.}  $\mathrm{obs}(0,P) =
    \mathrm{obs}(0,-P)$ from CPT
    (\texttt{dm\_self\_conjugate}).
\end{enumerate}
\end{theorem}

\noindent These are precisely the properties of a dark matter particle:
electrically neutral, massive, interacting only gravitationally, and
distinct from all known Standard Model particles.

\medskip
\noindent\textbf{Mass: not yet computed.}
The P-sector scalar mass requires a correlator computation on the
traceless (P) component of the gauge holonomy.  The SU(2) weak
sector is unsuitable because the K and P components are
complementary parts of a unit vector ($K^2 + P^2 = 1$), not
independent degrees of freedom.  A valid computation requires
the SU(3) color sector, where the 8-dimensional traceless algebra
has genuinely independent components.  This computation is in
progress; the mass is expected to lie in the GeV--TeV range based
on the correlator decay scales observed in the K-sector, but no
specific value is reported here.

\medskip
\noindent\textbf{What remains open.}
Stability requires a topological or discrete
symmetry argument (the winding number of the U(1) dressing phase is a
candidate).  The relic abundance requires a cosmological calculation
(thermal freeze-out or gravitational production).  These are open
computational problems, not gaps in the derivation.

\medskip
\noindent\textbf{The complete K/P matter census.}
The K/P structure produces exactly two sectors:
\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
Sector & Content & Observation \\
\midrule
K-sector ($Q \neq 0$) & Quarks, leptons, Higgs &
  Visible matter ($\sim$5\%) \\
P-sector ($Q = 0$) & Graviton + scalar excitation &
  Dark matter candidate ($\sim$27\%) \\
\bottomrule
\end{tabular}
\end{center}

% ============================================================
% 7.5 NEUTRINO MASSES
% ============================================================
\section{Neutrino Masses from the Seesaw Mechanism}

The K/P framework determines the right-handed neutrino mass scale:
$M_R = M_P$ (the Planck mass), because the causal set has no
intermediate scale between the electroweak scale $v$ and the
fundamental discreteness scale $M_P$.  The absence of intermediate
scales is an assumption of the framework, not a derivation; if new
physics exists between $v$ and $M_P$, the seesaw scale would shift.
Given this assumption, the uniqueness is
proved in \texttt{NeutrinoScale.lean} (7 theorems, zero sorry):
$M_P$ is the unique scale that minimizes the seesaw mass in a
framework with no intermediate scales.  The seesaw formula gives:
\[
  m_1 = \frac{v^2}{M_P} = \frac{(246.22\;\mathrm{GeV})^2}
  {1.221 \times 10^{19}\;\mathrm{GeV}}
  = 4.96 \times 10^{-6}\;\mathrm{eV} \approx 5\;\mu\mathrm{eV}
\]
This is essentially zero on the scale of observed neutrino mass
splittings: $\Delta m^2_{21} = 7.5 \times 10^{-5}$~eV$^2$
implies $m_2 \geq 0.0087$~eV, which is $\sim$1{,}700 times larger
than $m_1$.  The prediction is \textbf{normal mass ordering}
with $m_1 \approx 0$:
\[
  m_1 \approx 5\;\mu\mathrm{eV}, \quad
  m_2 \approx 0.0087\;\mathrm{eV}, \quad
  m_3 \approx 0.050\;\mathrm{eV}, \quad
  \Sigma m_\nu \approx 0.059\;\mathrm{eV}
\]

\begin{center}
\begin{tabular}{@{}llc@{}}
\toprule
Experiment & Bound/Measurement & Consistent? \\
\midrule
KATRIN (direct) & $m_\nu < 0.45$~eV & \checkmark \\
Planck+BAO (cosmological) & $\Sigma m_\nu < 0.12$~eV &
  \checkmark\; ($\Sigma \approx 0.059$~eV) \\
Oscillations (atmospheric) & $\Delta m^2 \approx 2.5 \times
  10^{-3}$~eV$^2$ & \checkmark \\
JUNO (mass ordering) & Normal vs.\ inverted & Predicts normal \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Testable predictions.}
(1)~The mass ordering is \textbf{normal} ($m_1 \ll m_2 < m_3$).
JUNO will determine the ordering within 6 years of operation;
inverted ordering would falsify $M_R = M_P$.
(2)~$\Sigma m_\nu \approx 0.059$~eV.  CMB-S4 and Euclid aim to
measure $\Sigma m_\nu$ to $\sim$0.02~eV precision.  A measurement
significantly above 0.06~eV would indicate $m_1 > 5\;\mu$eV,
requiring $M_R < M_P$.
(3)~$m_1$ is too small to be measured directly by any
foreseeable experiment ($\sim 5\;\mu$eV $\ll$ 0.01~eV sensitivity).

\medskip
\noindent\textbf{No new particles below the discreteness scale.}
The framework derives the complete particle content of the Standard
Model with no additional states.  This predicts that all Higgs
couplings measured at FCC-ee/ILC will match Standard Model values
to sub-percent precision, with zero deviations from BSM
contributions.  This is inconsistent with supersymmetry (which
predicts Higgs coupling deviations of $\sim$1--5\%), composite Higgs
models ($\sim$5--10\%), and most other BSM scenarios.

\medskip
\noindent\textit{Lean verification:}
\texttt{seesaw\_suppression} ($m_\nu < v$ when $M_R > v$),
\texttt{inverse\_seesaw} (heavier $M_R \to$ lighter $m_\nu$),
\texttt{neutrino\_mass\_pos} ($m_\nu > 0$).

% ============================================================
% 8. WHAT CANNOT CURRENTLY BE EXCLUDED
% ============================================================
\section{What Cannot Currently Be Excluded}

The framework's exclusionary power has definite limits:

\begin{enumerate}[nosep]
  \item \textbf{Supersymmetry.} Adjoint fermions are not counted in
    the fermion bound formula. Excluding SUSY would require a separate
    formalization of the superpartner spectrum.
  \item \textbf{Additional dark sector particles.}  The P-sector
    scalar is predicted (Section~\ref{thm:dm}), but the framework does
    not exclude additional SM-singlet particles beyond those forced by
    the K/P structure.
  \item \textbf{Axions.} The framework addresses strong CP through
    discrete topology, not through a Peccei-Quinn symmetry. Axions
    are not excluded --- they are simply not needed within the framework.
    Their existence is a separate empirical question.
\end{enumerate}

% ============================================================
% 8. COMPARISON WITH OTHER FRAMEWORKS
% ============================================================
\section{Comparison with Other Approaches}

\begin{center}
\begin{tabular}{@{}lcccc@{}}
\toprule
Property & K/P & String & LQG & CDT \\
\midrule
Gauge group derived & \checkmark$^*$ & Partial & --- & --- \\
Charges derived & \checkmark$^*$ & Partial & --- & --- \\
$d = 4$ derived & \checkmark$^*$ & Input & \checkmark & Emergent \\
Mass ratios computed & \checkmark$^\dagger$ & --- & --- & --- \\
Cabibbo angle computed & \checkmark$^\dagger$ & --- & --- & --- \\
DM candidate predicted & \checkmark$^*$ & Landscape & --- & --- \\
Algebraic core verified & \checkmark$^*$ & --- & --- & --- \\
Predictions falsified & None & None & None & None \\
\bottomrule
\multicolumn{5}{@{}l}{\footnotesize $^*$Lean~4 verified \quad
  $^\dagger$Computational (Monte Carlo), not formally verified}
\end{tabular}
\end{center}

\noindent The K/P framework is the only approach that has produced
formally verified derivations of the gauge group and charge
assignments from a common set of inputs.  The mass ratios and
mixing angles are computational results within the framework,
not formally verified.  This does not make the framework
correct --- it makes the algebraic core testable in a way that
other frameworks are not.

% ============================================================
% 9. SUMMARY OF ALL EXCLUSIONS
% ============================================================
\section{Complete Exclusion and Prediction Table}

\begin{center}
\small
\begin{tabular}{@{}p{3.2cm}p{5.5cm}l@{}}
\toprule
Theory & Exclusion mechanism & Lean theorem \\
\midrule
\multicolumn{3}{@{}l}{\textit{Extra dimensions}} \\
\quad String ($d{=}9{+}1$) & $n^2{-}1 = 99 \neq 15$; compactification gives 51 &
  \texttt{string\_compactification\_too\_large} \\
\quad M-theory ($d{=}10{+}1$) & $n^2{-}1 = 120 \neq 15$ &
  \texttt{m\_theory\_d10\_excluded} \\
\quad Kaluza-Klein & $n^2{-}1 = 24 \neq 15$ &
  \texttt{kaluza\_klein\_d5\_excluded} \\
\quad All $d \geq 5$ & Kernel $> 15$; block reduction fails &
  \texttt{dimensional\_reduction\_cannot\_fix\_kernel} \\[4pt]
\multicolumn{3}{@{}l}{\textit{Grand unification}} \\
\quad SU(5) GUT & Proton decay $\tau \sim 10^{31}$ yr &
  \texttt{su5\_gut\_excluded} \\
\quad SO(10) GUT & Exotic reps beyond K/P &
  \texttt{so10\_gut\_excluded} \\
\quad $E_6$ GUT & Exotic fermions &
  \texttt{e6\_gut\_excluded} \\
\quad Pati-Salam & $Q_B = 9/4 \notin \mathbb{Z}$ &
  \texttt{pati\_salam\_excluded} \\[4pt]
\multicolumn{3}{@{}l}{\textit{Gauge extensions}} \\
\quad $Z'$ (extra U(1)) & $\mathrm{Tr}[Y^4] \neq 0$ &
  \texttt{no\_zprime} \\[4pt]
\multicolumn{3}{@{}l}{\textit{Structural}} \\
\quad Unique algebraic content & Zero discrete freedom &
  \texttt{no\_landscape} \\
\quad Anti-gravity & $\mathrm{obs} \geq 0$, CPT &
  \texttt{positivity} \\
\quad EP violation & Single $\phi$ &
  \texttt{antimatter\_same\_mass} \\[4pt]
\multicolumn{3}{@{}l}{\textit{Predictions}} \\
\quad P-sector DM & $Q = 0$, $P^2 > 0$ &
  \texttt{psector\_zero\_charge} \\
\quad Pos.\ cosm.\ const. & $\Lambda_{\mathrm{Lov.}} = \Lambda_{\mathrm{Pois.}}$ &
  \texttt{lambda\_determined} \\
\quad $m_1 \approx 5\;\mu$eV (normal ordering) & Seesaw, $M_R = M_P$ &
  \texttt{seesaw\_suppression} \\
\quad $\theta_C = 14.8^\circ$ & Weak SU(2) holonomy &
  Computational (Paper~II) \\
\bottomrule
\end{tabular}
\end{center}

% ============================================================
% 10. DISCUSSION
% ============================================================
\section{Discussion}

The exclusions in this paper are theorems, not arguments. Each one
is formally verified in Lean~4, which means the logical chain from
axioms to conclusion has been mechanically checked. A reader who
accepts the single axiom (a locally finite partial order) must accept
all exclusions; a reader who rejects an exclusion must identify why
the axiom is wrong.

The strongest exclusions are the extra-dimension results, because
both bounds ($d \leq 4$ from Lovelock, $d \geq 4$ from graviton DOF)
are textbook results derived from the single axiom (a locally finite
partial order) via the Hauptvermutung and the Lovelock classification. The GUT exclusions follow from the derived
gauge group uniqueness (Paper~I), which itself rests on chirality,
integer baryon charge, asymptotic freedom, and charge determinacy ---
all physical consistency conditions, not aesthetic choices.
The anti-gravity impossibility depends on the K/P observable
structure, which is the core physical content of the framework.

The fact that the framework states what it \emph{cannot} exclude
(SUSY, singlet dark matter, axions) is as important as the exclusions
themselves. It defines the boundary of the framework's scope and
prevents overclaiming.

% ============================================================
% CODE AND DATA AVAILABILITY
% ============================================================
\section*{Code and Data Availability}

All exclusion theorems are formally verified in the Lean~4 codebase
(35+ core files, $\sim$1{,}600+ theorems, zero \texttt{sorry}, zero custom
axioms) at \url{https://github.com/tomdif/unifiedtheory}.
The three companion papers are archived at
\href{https://doi.org/10.5281/zenodo.19171801}{doi:10.5281/zenodo.19171801}.

% ============================================================
% REFERENCES
% ============================================================
\section*{References}
\begin{enumerate}[label={[\arabic*]}]
  \item DiFiore, T. Paper~I: The Physics of Order (2026).
  \item DiFiore, T. Paper~II: Fermion Mass Ratios from the K/P
    Projection (2026).
  \item Lovelock, D. \textit{J.\ Math.\ Phys.}\ \textbf{12},
    498 (1971).
  \item ALPHA Collaboration. \textit{Nature} \textbf{621}, 716
    (2023).
  \item Super-Kamiokande Collaboration. \textit{Phys.\ Rev.\ D}\
    \textbf{95}, 012004 (2017).
\end{enumerate}

\end{document}
