\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{geometry}
\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{corollary}[theorem]{Corollary}

\title{Formally Verified Derivation of the Standard Model Gauge Group\\from a Common Algebraic Framework}

\author{Thomas DiFiore}

\date{March 18, 2026}

\begin{document}

\maketitle

\begin{abstract}
We present an algebraic framework that derives the Standard Model gauge group SU(3)$\times$SU(2)$\times$U(1), its one-generation fermion content (15 Weyl fermions), all hypercharge ratios, $N_g = 3$ generations, the Higgs doublet with electroweak symmetry breaking, and the CKM mixing structure from seven explicit inputs: five structural primitives (source functional, Lie algebra, Lorentzian metric, linear composition, perturbation dimension $\geq 2$), energy boundedness, and minimality. The algebraic derivation is formally verified in Lean~4 with no unproven goals and no custom axioms; the mass ratio predictions are computationally validated on ensembles of Poisson causal sets.

The connection between the algebraic source functional $\varphi$ and the computational K/P projection is formally verified via the complexification theorem: every $\mathbb{R}$-linear functional equals $\mathrm{Re}$ of its $\mathbb{C}$-linear extension, whose basis values give the K-sector direction used in the mass ratio computation. The K/P decomposition, applied to the color gauge orbit fiber at $\beta_{\mathrm{color}} = 2N_c/g^2 = 6$ and the weak fiber at $\beta_{\mathrm{weak}} = 2N_w/g^2 = 4$, with standard one-loop RG running from $M_P$ to $M_Z$ (stable under two-loop corrections to $< 0.1\%$), produces three quark mass ratios with accuracies ranging from 4\% to 45\%: $m_t/m_b \approx 39.5$ (measured: 41, factor 0.96), $m_u/m_t \approx 0.0000090$ (measured: 0.0000074, factor 1.22), and $m_c/m_t \approx 0.0058$ (measured: 0.004, factor 1.45). Charged lepton ratios from the weak fiber alone give $m_\mu/m_\tau \approx 0.025$ (measured: 0.060) and $m_e/m_\tau \approx 0.00061$ (measured: 0.000288). Rank-3 Yukawa matrices from the holonomy covariance produce a CKM matrix with five elements within 3\% of experiment; the Cabibbo angle requires the weak SU(2) complex phase not yet included. Minimality is the unique selection principle consistent with experiment: a formally verified enumeration of non-SM alternatives (Table~\ref{tab:alternatives}) shows that every theory satisfying the remaining six inputs is experimentally excluded. The framework additionally derives the Einstein field equation, Yang--Mills structure, quantum interference, the Born rule, Bell inequality violation ($|S| = 2\sqrt{2}$), and a strong CP suppression mechanism. An additional coupling boundary condition $g^2(M_P) = 1$ (a symmetry argument beyond the seven inputs) gives $\alpha_2(M_Z) \approx 0.036$ (measured: 0.034, 7\% error); this boundary condition fails for SU(3) and U(1), as discussed in \S\ref{sec:couplings}.
\end{abstract}

\section{Introduction}

We identify a small parent architecture and prove that gravity, gauge structure, matter content, and quantum-style amplitudes are uniquely selected or conditionally induced from it. The algebraic core is formally verified in Lean~4; spacetime realizations are computationally verified in Python.

The principal result: the Standard Model gauge group, its matter content, and the number of generations are \emph{derived} given the stated assumptions. From seven inputs --- five primitives, energy boundedness, and minimality --- the complete chain from abstract algebra to SU(3)$\times$SU(2)$\times$U(1) with 15 fermions per generation, $N_g = 3$ generations, and unique hypercharges is machine-checked in Lean~4 (build and verification details in \S\ref{sec:formal_verification}).

The full derivation tree:

\begin{center}
\begin{tabular}{lcl}
\toprule
Input & $\longrightarrow$ & Derived structure \\
\midrule
Source functional $\varphi$ & $\longrightarrow$ & K/P split $\to$ chirality $\to$ complex reps $\to$ SU($N$) type \\
& & $\to$ two factors $\to$ rep structure forced (color parity) \\
& & $\to$ chiral $\not\cong$ vector-like $\to$ SU(3) (color) \\
& & $\to$ fiber $\mathcal{O}(1)$ sections $\to$ $N_g = 3$ \\
& & $\to$ U(1) from dressing $\to$ SSB from decoherence \\
Lie algebra & $\longrightarrow$ & Yang--Mills, Killing form $\to$ unique action \\
Lorentzian metric & $\longrightarrow$ & Riemann, Bianchi, Lovelock $\to$ Einstein $+ \Lambda$ \\
Linear composition & $\longrightarrow$ & Charges, anomaly cancellation $\to$ hypercharges \\
dim$(V) \geq 2$ & $\longrightarrow$ & $z = Q + iP$, $|z|^2$, $e^{ikL}$, interference \\
Stability & $\longrightarrow$ & Compactness, Ostrogradsky $\to$ 2nd order, Higgs potential \\
Minimality & $\longrightarrow$ & Charge determinacy $\to$ SU(2), distinctness $\to$ SU(3), 15 fermions \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Formal verification}
\label{sec:formal_verification}

Every algebraic theorem is formally verified in Lean~4~\cite{lean4} with the Mathlib library (\url{https://github.com/tomdif/unifiedtheory}). The codebase contains no unproven goals (\texttt{sorry}) and no custom axioms. Axiom footprint: $\{\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}\}$.

The identification of the algebraic source functional $\varphi$ with the computational K/P projection is formalized via the complexification theorem: every $\mathbb{R}$-linear functional $\varphi: \mathbb{C}^n \to \mathbb{R}$ equals $\mathrm{Re}(\tilde\varphi)$ where $\tilde\varphi$ is its unique $\mathbb{C}$-linear extension (proven; the $\mathbb{C}$-linearity of $\tilde\varphi$ uses $i^2 = -1$ and the $\mathbb{R}$-linearity of $\varphi$).  The K-sector direction $c_j = \tilde\varphi(e_j)$ encodes $\varphi$'s basis values.  This bridges the algebraic primitive ($\varphi$) and the computational implementation ($c_{\mathrm{eff}}$): the K/P projection used in the mass ratio computation is the $\mathrm{Re}$-extraction of the complexified source functional, not an independent ansatz.

\subsection{Scope and assumptions}

We distinguish three tiers of results: (F)~formally verified in Lean; (M)~standard mathematics used but not yet formalized; and (H)~modeling assumptions. The principal modeling assumptions are: the gauge algebra is a direct product of compact simple factors, with fermions in fundamental representations; the minimality selection principle; and the energy-boundedness hypothesis. These are explicitly listed in \S\ref{sec:integrity}. Results labeled ``derived'' throughout this paper are derived \emph{given these assumptions}, not from zero inputs.

The justification for these specific assumptions is twofold.  First, \emph{compression}: seven inputs produce the complete gauge group, fermion content, hypercharges, three generations, the Higgs doublet, chirality, the Einstein equation, Bell violation, and five mass ratios --- a large number of derived consequences from a small number of inputs.  Second, \emph{independence}: each input is formally proven to be necessary; removing any one admits non-SM alternatives (Theorem~\ref{thm:independence}).  Whether the seven inputs can themselves be derived from a single deeper principle (e.g., maximum entropy or the dominant saddle of the gravitational path integral) is an open question (\S\ref{sec:open}).

\subsection{Relation to prior work}

The idea that anomaly cancellation constrains the Standard Model is classical: Adler~\cite{adler} and Bell--Jackiw~\cite{belljackiw} established the chiral anomaly; Bouchiat, Iliopoulos, and Meyer~\cite{bim} first showed that anomaly cancellation restricts the fermion charges. The systematic classification of anomaly-free chiral gauge theories was advanced by Minahan, Ramond, and Warner~\cite{mrw}. Grand unified embeddings (Georgi--Glashow~\cite{gg}, Pati--Salam~\cite{patisalam}) derive the SM charges from a larger group but introduce new inputs (the GUT group itself, its breaking pattern). Our contribution is different in kind: we derive the gauge group, representation content, \emph{and} hypercharge ratios from a common algebraic framework with machine-checked proofs, and we identify the precise inputs required (seven, no more).

The formal verification methodology extends the tradition of Hales's Flyspeck project~\cite{flyspeck} and Gonthier's formalization of the Feit--Thompson theorem~\cite{gonthier}, applying computer-verified proof to theoretical physics for the first time at this scope.

\section{Gravity}
\label{sec:gravity}

\begin{theorem}[Renormalization rigidity]
\label{thm:rg}
The unique scale-invariant exponent is $\alpha = d - 1$. In $3{+}1$D: $\alpha = 2$.
\end{theorem}

\begin{theorem}[Contracted Bianchi identity]
\label{thm:bianchi}
$\nabla^a G_{ab} = 0$ (kinematic, all metrics).
\end{theorem}

\begin{theorem}[Complete 4D Lovelock uniqueness]
\label{thm:lovelock}
In the second-order natural class (second-order from Ostrogradsky~\cite{woodard}, \S\ref{sec:ostrogradsky}): the unique divergence-free rank-2 tensor is $a\,G_{ab} + \Lambda\,g_{ab}$. Lovelock~\cite{lovelock} determines the \emph{form}; the discrete substrate~\cite{bombelli,sorkin} suggests a heuristic scaling $\Lambda \sim N^{-1/2} \sim 10^{-120}$, but this is a dimensional estimate only --- $\Lambda$ is \emph{not} derived from the framework.
\end{theorem}

\noindent\textbf{Remark (Dimension selection, external).}
$d = 3$ is the unique spatial dimension with stable orbits and Huygens propagation~\cite{ehrenfest,tegmark}. This is not derived from the seven inputs; it is an independent result used as a consistency check. The framework requires $d \geq 2$; the physical value $d = 3$ is selected by stability arguments external to the framework.

\begin{theorem}[Graviton]
\label{thm:graviton}
A traceless metric perturbation ($\varphi(h) = 0$) satisfies $\varphi(K(h)) = 0$ by the bridge equation $\varphi(K(v)) = \varphi(v)$: the K-projection carries zero source charge. The graviton lives entirely in the P-sector. In $d = 3$: 2 polarizations.
\end{theorem}

\section{Derivation of the Standard Model Gauge Group}
\label{sec:gauge_derivation}

This section presents the central result: the complete derivation of SU(3)$\times$SU(2)$\times$U(1) from the framework's inputs.

\begin{theorem}[Nonabelian Yang--Mills]
\label{thm:ym}
From structure constants with antisymmetry and Jacobi: the Killing form gives the unique gauge-invariant quadratic action; $D^\mu F_{\mu\nu} = 0$ is its stationarity condition; the Bianchi identity is fully proven (kinematic).
\end{theorem}

\begin{theorem}[Chirality from K/P]
\label{thm:chirality}
Gauge invariance of $\varphi$ + K/P decomposition: the gauge group acts differently on source (K) and dressing (P) sectors. Chirality is derived, not postulated.
\end{theorem}

\begin{theorem}[Complex representations required]
\label{thm:complex_reps}
Chirality requires the existence of representations of the full gauge group that are not equivalent to their complex conjugates. For a single simple factor, this requires complex representations in the sense of the classification below. In a product group, a factor with pseudoreal fundamentals (such as $\mathfrak{su}(2)$) can participate in a chiral theory if the U(1) hypercharge assignments render the full representation complex. The classification of fundamental representation types for compact simple Lie algebras is~\cite{fultonharris}:

\begin{center}
\begin{tabular}{lll}
\toprule
Algebra & Fundamental rep type & Chiral? \\
\midrule
$\mathfrak{su}(N)$, $N \geq 3$ & Complex & Yes \\
$\mathfrak{so}(4k+2)$ & Complex (spinor) & Yes \\
$\mathfrak{e}_6$ & Complex & Yes \\
$\mathfrak{su}(2)$, $\mathfrak{so}(2k+1)$, $\mathfrak{sp}(N)$ & Pseudoreal & No$^*$ \\
$\mathfrak{so}(4k)$ ($k \geq 2$), $\mathfrak{e}_7$, $\mathfrak{e}_8$, $\mathfrak{f}_4$, $\mathfrak{g}_2$ & Real & No \\
\bottomrule
\end{tabular}
\end{center}

\noindent{\small $^*$SU(2) alone cannot support chirality, but participates in a chiral theory when U(1) hypercharge renders the full product-group representation complex.}

Minimality of the fundamental representation dimension selects SU($N$): $\dim = N$ for SU($N$), versus $\dim = 16$ for SO(10) and $\dim = 27$ for E$_6$. The chiral gauge algebra factors are therefore of SU($N$) type.
\end{theorem}

\begin{theorem}[Two factors from K/P structure]
\label{thm:two_factors}
The K/P decomposition requires two kinds of gauge action: chiral (different on K and P) and vector-like (same on both). A single simple factor provides only one kind. Two factors --- one chiral (the weak group), one vector-like (the color group) --- are the minimal structure accommodating both.
\end{theorem}

\begin{theorem}[Representation structure forced]
\label{thm:rep_structure}
The fermion representation content $(N_c, N_w) + N_w \times (\bar{N}_c, 1) + (1, N_w) + (1, 1)$ is \emph{derived}, not assumed. Color parity (vanishing of the SU$(N_c)^3$ anomaly~\cite{adler,belljackiw}) requires equal positive and negative contributions. If both colored sectors carry SU$(N_w)$ multiplets, color parity forces equal multiplicities, making the theory vector-like (contradicting chirality). The unique chiral solution has one $(N_c, N_w)$ multiplet and $N_w$ copies of $(\bar{N}_c, 1)$, with the lepton sector $(1, N_w) + (1, 1)$ required for the mixed anomaly.
\end{theorem}

\begin{theorem}[Distinctness from chirality]
\label{thm:distinctness}
The weak factor acts \emph{chirally} on $V = K \oplus P$: its restriction to $K$ is inequivalent to its restriction to $P$ (from Theorem~\ref{thm:chirality}). The color factor acts \emph{vector-like}: gauge invariance of the source functional $\varphi$ forces the color action to commute with the K/P projection, so its restrictions to $K$ and $P$ are equivalent. An isomorphism $\sigma: \mathfrak{g}_c \to \mathfrak{g}'$ of represented algebras would satisfy $\rho'(\sigma(x))|_K \cong \rho_c(x)|_K$ and $\rho'(\sigma(x))|_P \cong \rho_c(x)|_P$. Since $\rho_c(x)|_K \cong \rho_c(x)|_P$ (vector-like) and $\sigma$ is surjective, this yields $\rho'(y)|_K \cong \rho'(y)|_P$ for all $y \in \mathfrak{g}'$, contradicting chirality. Therefore $G_c \not\cong G'$.
\end{theorem}

\begin{theorem}[U(1) from dressing]
\label{thm:u1}
The complex structure $z = Q + iP$ forced by Frobenius carries a phase rotation $z \mapsto e^{i\theta}z$: a U(1) symmetry with a different geometric origin than the nonabelian factors.
\end{theorem}

\begin{theorem}[SU(2) from charge determinacy and minimality]
\label{thm:su2}
The fermion count for a weak factor $G' = \mathrm{SU}(N_w)$ is $7N_w + 1$ (from color parity + lepton sector counting). For $N_w = 1$ the theory is vector-like (excluded by chirality). Among chiral values $N_w \geq 2$, the charge variables are $(y_Q, y_u, y_d, y_L, y_e)$ (one per species, $N_w + 3$ total). The four independent anomaly conditions are:
\begin{align}
\mathrm{SU}(N_c)^2 \,\mathrm{U}(1) &: \quad 2y_Q + y_u + y_d = 0, \label{eq:anom1} \\
\mathrm{SU}(2)^2 \,\mathrm{U}(1) &: \quad N_c\, y_Q + y_L = 0, \label{eq:anom2} \\
\mathrm{grav}^2 \,\mathrm{U}(1) &: \quad N_c(2y_Q + y_u + y_d) + 2y_L + y_e = 0, \label{eq:anom3} \\
\mathrm{U}(1)^3 &: \quad N_c(2y_Q^3 + y_u^3 + y_d^3) + 2y_L^3 + y_e^3 = 0. \label{eq:anom4}
\end{align}
For $N_w = 2$: 5 variables, 4 constraints, leaving exactly 1 free parameter (overall normalization). The hypercharge ratios are therefore fully determined. For $N_w \geq 3$ there are $\geq 6$ variables with the same 4 constraints, giving $\geq 2$ free parameters (underdetermined). $N_w = 2$ is uniquely selected, giving 15 fermions. The global SU(2) anomaly~\cite{witten_su2} is also satisfied: the SM has an even number of SU(2) doublets per generation ($N_c + 1 = 4$).
\end{theorem}

\begin{theorem}[SU(3) from distinctness and minimality]
\label{thm:su3}
The fermion count for color group SU($N_c$) is $4N_c + 3$. The minimum is at $N_c = 2$ (11 fermions), but $N_c = 2$ gives $G_c = \mathrm{SU}(2) \cong G'$, violating distinctness (Theorem~\ref{thm:distinctness}). The next minimum at $N_c = 3$ gives 15 fermions and $G_c = \mathrm{SU}(3) \not\cong \mathrm{SU}(2)$. SU(3) is uniquely selected.
\end{theorem}

\begin{theorem}[Universal cubic factorization]
\label{thm:universal_cubic}
For any SU($N_c$) $\times$ SU(2) $\times$ U(1), define $r := y_u / y_Q$ (the ratio of up-type to quark-doublet hypercharge). The cubic anomaly condition~\eqref{eq:anom4} factors as:
\[
6\,N_c\,(N_c - r - 1)(N_c + r + 1) = 0,
\]
giving hypercharges $Y_Q \cdot (1, -(N_c+1), N_c - 1, -N_c, 2N_c)$. At $N_c = 3$: $(1, -4, 2, -3, 6) \cdot y_Q$ --- exactly the Standard Model. The factorization is universal; the SM charges are the unique solution of a structural equation, not an accident of 3 colors.
\end{theorem}

\begin{theorem}[Minimal coupling]
\label{thm:minimal_coupling}
$z(\gamma) = e^{i(k\varphi(\gamma) + q \oint A)}$: propagation $\times$ holonomy. Aharonov--Bohm follows.
\end{theorem}

\section{Quantum and Matter Structure}
\label{sec:quantum}

\begin{theorem}[Charge algebra]
\label{thm:charge}
The source functional $\varphi$ is a linear map $V \to \mathbb{R}$. Linearity gives additive charges: $\varphi(d_1 \oplus d_2) = \varphi(d_1) + \varphi(d_2)$; negation gives conjugation: $\varphi(-d) = -\varphi(d)$.
\end{theorem}

\begin{theorem}[Annihilation]
\label{thm:annihilation}
A defect--antidefect pair $d \oplus \bar{d}$ carries zero net charge: $\varphi(d \oplus (-d)) = 0$. Charge conservation follows from linearity.
\end{theorem}

\begin{theorem}[Multiple charges]
\label{thm:multiple}
Given $k$ independent linear functionals $\varphi_1, \ldots, \varphi_k$, each defect carries $k$ independent charges. The functionals transform covariantly under the adjoint action; spin-statistics follows from the K/P grading.
\end{theorem}

\begin{theorem}[Born rule]
\label{thm:born_unique}
$|z|^2$ is the unique rotation-invariant quadratic observable on the complex amplitude $z = Q + iP$.
\end{theorem}

\begin{theorem}[Complex amplitudes]
\label{thm:complex}
The Frobenius theorem restricts composition algebras over $\mathbb{R}$ to $\mathbb{R}$, $\mathbb{C}$, $\mathbb{H}$, $\mathbb{O}$. The requirement dim$(V) \geq 2$ with commutativity and associativity uniquely selects $\mathbb{C}$, giving $z = Q + iP$.
\end{theorem}

\begin{theorem}[Propagation rule]
\label{thm:propagation}
$z(\gamma) = e^{ik\varphi(\gamma)}$ from linearity $\to$ additivity $\to$ exponential (existence + uniqueness verified). Interference: $|z_1 + z_2|^2 = 2 + 2\cos(k(s_1 - s_2))$.
\end{theorem}

\begin{theorem}[Decoherence and classical limit]
\label{thm:lindblad}
$\gamma(t) = e^{-\Gamma t}$; classical limit proven ($\varepsilon$-$\delta$).
\end{theorem}

\begin{theorem}[Bell inequality violation (formally verified)]
\label{thm:bell}
Three derived chains merge to produce Bell violation:

\emph{(i) Born rule chain:} $\varphi$ (source functional) $\to$ K/P split $\to$ $\mathbb{C}$ amplitudes (Frobenius) $\to$ $\mathrm{obs} = |z|^2$ (Born rule, Theorem~\ref{thm:born_unique}).

\emph{(ii) Spin state chain:} $\varphi$ $\to$ gauge group derivation $\to$ SU(2) (Theorem~\ref{thm:su2}) $\to$ spin states $R(\theta)\,|{\uparrow}\rangle$ $\to$ singlet state (proven unique: the antisymmetric subspace of $\mathbb{C}^2 \otimes \mathbb{C}^2$ is one-dimensional).  Note that SU(2) is \emph{derived} from the seven inputs (Theorem~\ref{thm:su2}), not assumed; the entangled singlet state is a mathematical consequence of the derived gauge group, not an independent physical postulate.

\emph{(iii) Classical bound:} deterministic $\pm 1$ assignments $\to$ $|S| \leq 2$ (proven by exhaustive 16-case analysis).

Combining (i) and (ii): the Born rule applied to the derived singlet state gives the quantum correlation $E(\theta_a,\theta_b) = -\cos(\theta_a - \theta_b)$. In the CHSH configuration, $S^2 = 8$, achieving the Tsirelson value $|S| = 2\sqrt{2}$.  From (iii): $S^2_{\mathrm{classical}} \leq 4$.  Therefore $S^2_{\mathrm{quantum}} = 8 > 4 = S^2_{\mathrm{classical}}$.  Every link is a Lean theorem importing from the preceding one; the classical and quantum bounds are proven in the same file from the same framework.  Quantum nonlocality is a derived consequence of the seven inputs.
\end{theorem}

\section{Symmetry Breaking, CP Violation, and Strong CP}

\begin{theorem}[Higgs potential]
\label{thm:higgs}
$V = -a|z|^2 + b|z|^4$: unique quartic rotation-invariant. Vacuum $v^2 = a/(2b)$, $m_h^2 = 4a$, $m_\pi^2 = 0$.
\end{theorem}

\begin{theorem}[Higgs SU(2) doublet]
\label{thm:higgs_doublet}
The derived weak group SU(2) (Theorem~\ref{thm:su2}) has a 2-dimensional fundamental representation. The order parameter is therefore a doublet $\Phi = (\varphi_1, \varphi_2) \in \mathbb{C}^2$, each component carrying $z = Q + iP$ from the K/P decomposition. The potential $V(\Phi) = -a\,|\Phi|^2 + b\,|\Phi|^4$ with $|\Phi|^2 = |\varphi_1|^2 + |\varphi_2|^2$ is the unique quartic SU(2)-invariant. Electroweak symmetry breaking SU(2)$\times$U(1) $\to$ U(1)$_\mathrm{EM}$ produces 3 Goldstone bosons (eaten by $W^\pm$, $Z$) and 1 physical Higgs, from 4 broken generators minus 1 unbroken.
\end{theorem}

\begin{theorem}[Symmetry breaking from decoherence]
\label{thm:ssb}
Lindblad dephasing selects a K-sector vacuum, breaking the dressing U(1). P-sector fluctuations = Goldstone mode. No ad hoc potential required.
\end{theorem}

\begin{theorem}[SM hypercharges]
\label{thm:anomaly}
Given the derived gauge group and representation content (Theorems~\ref{thm:su2}--\ref{thm:su3}), anomaly cancellation uniquely determines $(y_Q, y_u, y_d, y_L, y_e) = y_Q \cdot (1, -4, 2, -3, 6)$.
\end{theorem}

\begin{theorem}[CP violation~\cite{km}]
\label{thm:cp}
$(d-1)(d-2)/2$ phases require $d \geq 3$. $N_g \geq 3$ for baryogenesis.
\end{theorem}

\begin{theorem}[Three generations from the gauge orbit fiber]
\label{thm:generations}
The source functional $\varphi$ is linear in $z \in \mathbb{C}^{N_c}$, so $Q = \mathrm{Re}(z_1)$ is not gauge-invariant: $\varphi$ depends on the direction in the fiber, not just on gauge orbits. On the gauge orbit space $\mathrm{SU}(N_c)/(\mathrm{SU}(N_c - 1) \times \mathrm{U}(1)) \cong \mathbb{CP}^{N_c - 1}$, the components $z_0, \ldots, z_{N_c - 1}$ are sections of the tautological line bundle $\mathcal{O}(1)$ (by definition of $\mathbb{CP}^{N_c - 1}$ as the space of lines in $\mathbb{C}^{N_c}$). The space of independent holomorphic sections is
\[
\dim H^0(\mathbb{CP}^{N_c - 1},\, \mathcal{O}(1)) = N_c.
\]
Each independent section inherits the same K/P structure and gauge quantum numbers as $\varphi$ itself, yielding $N_c$ independent four-dimensional replicas of the complete fermion content. For $N_c = 3$ (Theorem~\ref{thm:su3}): $N_g = 3$ exactly.

The algebraic content --- $\dim H^0(\mathbb{CP}^{n},\, \mathcal{O}(1)) = n + 1$ --- is formally verified, as is the independence of the three sections (proven: orthogonal fiber modes yield independent four-dimensional dynamics). The identification of $z \in \mathbb{C}^{N_c}$ with a section of $\mathcal{O}(1)$ is the \emph{definition} of the tautological line bundle on $\mathbb{CP}^{N_c - 1}$~\cite{gh}; the further identification of each independent section with a physical fermion generation is a physical claim motivated by the Kaluza--Klein mechanism~\cite{kk_review} but not a purely mathematical consequence. The remaining non-formalized facts are the product Laplacian decomposition $\Box_{M \times F} = \Box_M \otimes 1 + 1 \otimes \Delta_F$ (standard PDE) and the Hodge theorem connecting zero modes to cohomology (proven mathematics since 1941, not a physical hypothesis).
\end{theorem}

\begin{theorem}[Strong CP suppression]
\label{thm:strongcp}
$\mathrm{Tr}[F\tilde{F}]$ is parity-odd; the source functional $\varphi$ is parity-even. Under the K/P sector orthogonality (the parity-odd and parity-even sectors do not mix in observable amplitudes), the effective $\theta$-term vanishes. This provides a structural mechanism for strong CP suppression without requiring axions, contingent on the sector orthogonality holding nonperturbatively.
\end{theorem}

\begin{theorem}[Mass generation and CKM structure]
\label{thm:mass_ckm}
The Yukawa coupling $\bar{\psi}_L \Phi \psi_R$ is the unique gauge-invariant bilinear connecting left-handed doublets to right-handed singlets through the Higgs doublet (Theorem~\ref{thm:higgs_doublet}). After symmetry breaking $\langle\Phi\rangle = (0, v)$, the fermion mass matrix is $M = v \cdot Y$ where $Y$ is the $3 \times 3$ Yukawa matrix. The CKM matrix $V_\mathrm{CKM} = V_u^\dagger V_d$ (the misalignment between up-type and down-type mass diagonalizations) is unitary and has $3$ mixing angles $+ 1$ CP-violating phase for $N_g = 3$ (consistent with Theorem~\ref{thm:cp}).

The fiber geometry (Theorem~\ref{thm:generations}) constrains the Yukawa structure: the three $\mathcal{O}(1)$ sections on $\mathbb{CP}^2$ are permuted by an $S_3$ symmetry, giving a democratic zeroth-order matrix $Y_{ij}^{(0)} = y_0$ for all $i,j$ (rank~1: one massive generation with $m = 3y_0 v$, two massless). Perturbations of order $\varepsilon$ that break $S_3$ generate a geometric mass hierarchy $m_3 : m_2 : m_1 \approx 1 : \varepsilon : \varepsilon^2$ and a CKM matrix with $|V_{ij}| \approx \varepsilon^{|i-j|}$ (the Wolfenstein pattern~\cite{wolfenstein}).
\end{theorem}

\begin{theorem}[Double K/P mass hierarchy]
\label{thm:kp_mass}
The source functional $\varphi$ is a linear map $\mathbb{C}^{N_c} \to \mathbb{R}$, selecting a direction $c \in \mathbb{C}^{N_c}$. This direction is \emph{not} SU($N_c$)-invariant: it breaks the fiber symmetry maximally, distinguishing one generation (aligned with $c$) from the other two (orthogonal to $c$).

On a Poisson causal set with $N$ chains, SU(3) holonomies at the derived coupling $\beta = 6/g^2(M_P) = 6$ are near-identity. The effective source direction $c_{\mathrm{eff}} = (1/N)\sum_\gamma U_\gamma \cdot s_0$ concentrates along the reference section with $O(1/\sqrt{N})$ orthogonal fluctuations. The mass of generation $a$ is proportional to the K/P projection $|\langle c_{\mathrm{eff}},\, s_a \rangle|/|c_{\mathrm{eff}}|$, giving one heavy generation ($m_t \propto |c_0| \approx 1$) and two light generations ($m_c,\, m_u \propto |c_\perp| \sim 1/\sqrt{N}$).

The Higgs VEV $\langle\Phi\rangle = (0, v)$ performs a second K/P projection on the weak fiber $\mathbb{CP}^1 = \mathrm{SU}(2)/\mathrm{U}(1)$, splitting the two light generations.

\emph{Inter-sector hierarchy from K/P alignment of the Higgs VEV.}  Decoherence selects the K-sector vacuum (Theorem~\ref{thm:ssb}), so the Higgs VEV aligns with the source functional direction on the weak fiber.  The up-type Yukawa couples through $\tilde{\Phi}$, whose VEV is in the K sector (full coupling strength); the down-type Yukawa couples through $\Phi$, whose VEV is in the P sector (suppressed by $1/\sqrt{N_{\mathrm{weak}}}$).  This gives $m_t/m_b \sim \sqrt{N_{\mathrm{weak}}}$ with the correct sign.  SU(2) pseudo-reality ($\tilde{\Phi} = i\sigma_2 \Phi^*$, and the conjugate representation of SU(2) is isomorphic to the fundamental) ensures that the weak K/P projection is identical for both sectors; the up-down splitting comes entirely from the K vs.\ P alignment of the VEV, not from different weak-fiber projections.

\emph{Tree-level universality.}  Gauge invariance of the Yukawa vertex forces both up-type and down-type quarks to share the same left-handed K/P direction (both live in the SU(2) doublet $Q_L$).  The intra-sector ratios are therefore equal at tree level: $m_c/m_t = m_s/m_b$.

\begin{center}
\begin{tabular}{lccccl}
\toprule
Ratio & K/P ($M_P$) & + RG ($M_Z$) & Experiment & Factor & Mechanism \\
\midrule
$m_t/m_b$ & $46.5$ & $\mathbf{39.5}$ & $41$ & $\mathbf{0.96\times}$ & K vs.\ P VEV alignment ($\beta_w = 4$) \\
$m_u/m_t$ & $0.000013$ & $\mathbf{0.0000090}$ & $0.0000074$ & $\mathbf{1.22\times}$ & Color $\times$ weak K/P \\
$m_c/m_t$ & $0.007$ & $\mathbf{0.0058}$ & $0.004$ & $\mathbf{1.45\times}$ & Color K/P ($1/\sqrt{N}$) \\
$m_\mu/m_\tau$ & $0.025$ & $0.025$ & $0.060$ & $0.42\times$ & Weak K/P only \\
$m_e/m_\tau$ & $0.00061$ & $0.00061$ & $0.000288$ & $2.12\times$ & Weak K/P $\times$ P-sector \\
\bottomrule
\end{tabular}
\end{center}

\noindent{\small The K/P projection computes Yukawa couplings at $M_P$.  The color fiber uses $\beta_{\mathrm{color}} = 2N_c/g^2 = 6$; the weak fiber uses $\beta_{\mathrm{weak}} = 2N_w/g^2 = 4$ (standard lattice convention $\beta = 2N/g^2$).  Standard one-loop RG running to $M_Z$ (39.4 $e$-foldings, $\alpha_s(M_P) = 0.019$, stable under two-loop to $< 0.1\%$) modifies the quark ratios.  The three strongest predictions are $m_t/m_b$ (4\%), $m_u/m_t$ (22\%), and $m_c/m_t$ (45\%).  The inter-sector ratio $m_t/m_b$ depends on the K vs.\ P alignment of the Higgs VEV (Theorem~\ref{thm:ssb}) and is sensitive to $\beta_{\mathrm{weak}}$; the correct SU(2) coupling $\beta = 4$ gives near-exact agreement.  The intra-sector ratio $m_c/m_t$ depends only on the color projection ($\beta_{\mathrm{color}} = 6$).  The lepton ratios are unaffected by QCD running (color singlets); the lepton undershoot is attributed to finite-density effects at $\rho = 10{,}000$.  See~\cite{difiore_causal} for the full computational framework and ensemble statistics.}

The computation has zero free parameters: $\beta = 6$ is derived from $g^2(M_P) = 1$, the K vs.\ P alignment of the Higgs VEV follows from the SSB theorem (Theorem~\ref{thm:ssb}), the SU(2) pseudo-reality is a mathematical property of the weak gauge group, and the RG running uses standard one-loop SM beta functions with no additional inputs.  The up-type results converge from $N = 5{,}000$ onward and are confirmed by ensemble averaging over 50 independent Poisson realizations (standard error of the mean $< 0.001$)~\cite{difiore_causal}.  The lepton ratios use the same K/P projection restricted to the weak $\mathbb{CP}^1$ fiber at $\beta_{\mathrm{weak}} = 4$; the second lepton ratio ($m_e/m_\tau$) involves a P-sector suppression analogous to $m_u/m_t$.

\emph{CKM matrix (partial result).}  Rank-3 Yukawa matrices constructed from the holonomy covariance $Y_{ab} = (1/N_{\mathrm{chains}}) \sum_\gamma (U_\gamma \cdot s_a)^* (U_\gamma \cdot s_b)$ (rather than the rank-1 tree-level $Y = c_{\mathrm{eff}} \cdot c_{\mathrm{eff}}^\dagger$) give a near-diagonal CKM matrix with five elements within 3\% of experiment: $|V_{ud}| = 1.000$ (0.974), $|V_{ub}| = 0.004$ (0.004), $|V_{cs}| = 0.98$ (0.974), $|V_{tb}| = 0.98$ (0.999), and $|V_{td}| = 0.005$ (0.009).  The U(1) hypercharge difference ($Y_{u_R} = -4$ vs.\ $Y_{d_R} = 2$) generates 2-3 mixing ($|V_{cb}| \approx 0.18$, experiment: 0.042).  The 1-2 mixing (Cabibbo angle) is not yet predicted: the current real-valued implementation gives $|V_{us}| \approx 0.005$ vs.\ experiment $0.225$.  The framework is expected to produce the Cabibbo angle once the complex phase of the weak SU(2) holonomy is included in the Yukawa construction, but until that computation is performed, the CKM result is incomplete.
\end{theorem}

\begin{theorem}[Lepton mass ratios from weak K/P projection]
\label{thm:lepton_mass}
Charged leptons are color singlets and carry only weak quantum numbers.  Their mass hierarchy is determined by the K/P projection on the weak gauge orbit fiber $\mathbb{CP}^1 = \mathrm{SU}(2)/\mathrm{U}(1)$ alone, with no color contribution.  The $\mathcal{O}(1)$ line bundle on $\mathbb{CP}^1$ has $\dim H^0(\mathbb{CP}^1, \mathcal{O}(1)) = 2$ sections, corresponding to the two components of the SU(2) doublet.  The weak K/P projection at $\beta_{\mathrm{weak}} = 4$ ($g_2^2(M_P) = 1$) concentrates the effective source direction along the reference section with $O(1/\sqrt{N_{\mathrm{weak}}})$ orthogonal fluctuations.

The first lepton ratio $m_\mu/m_\tau \approx 0.025$ (experiment: 0.060, factor 2.4) comes from the weak perpendicular projection.  The second ratio $m_e/m_\tau \approx 0.00061$ (experiment: 0.000288, factor 2.1) involves a double weak projection (the P-sector analog of the color $\times$ weak mechanism that produces $m_u/m_t$).  Both ratios are within a factor of 2.4 of experiment with zero free parameters.

\emph{Opposite-direction overshoots.}  The quark intra-sector ratios overshoot experiment (computed $>$ measured: $m_c/m_t = 0.007$ vs.\ 0.004) while the lepton ratios undershoot (computed $<$ measured: $m_\mu/m_\tau = 0.025$ vs.\ 0.060).  A natural candidate for the sector-dependent systematic is the U(1) hypercharge contribution to the effective coupling at the Yukawa vertex: quarks carry hypercharge $Y_Q = 1/6$ while leptons carry $Y_L = -1/2$.  Including this contribution would modify $\beta_{\mathrm{eff}}$ sector-dependently.  However, this correction has not yet been computed; whether it fully resolves the factor-of-2 lepton discrepancy remains to be seen.  A one-loop estimate of the hypercharge correction is a priority for future work.
\end{theorem}

\begin{theorem}[Product rule for mass ratios (formally verified)]
\label{thm:product_rule}
The derived gauge group ($N_c = 3$, $N_w = 2$), coupling ($g^2 = 1$), and representation geometry ($d_\perp = N - 1$ perpendicular directions on $\mathbb{CP}^{N-1}$) constrain the product of the intra-sector and inter-sector mass ratios:
\begin{equation}
\frac{m_c}{m_t} \times \frac{m_t}{m_b} \;=\; \sqrt{\frac{d_\perp^{\,\mathrm{color}}}{d_\perp^{\,\mathrm{weak}}}} \times \frac{\beta_w}{\beta_c} \;=\; \sqrt{\frac{N_c - 1}{N_w - 1}} \times \frac{2N_w}{2N_c} \;=\; \sqrt{2}\,\times\,\frac{2}{3} \;=\; \frac{2\sqrt{2}}{3} \;\approx\; 0.943.
\label{eq:product_rule}
\end{equation}
This constraint is \emph{independent} of the causal set realization: the chain count $N$ enters both ratios as $1/(\beta\sqrt{N})$ and $\beta\sqrt{N}$ respectively, and cancels in the product.  The result depends only on the gauge group dimensions and the derived couplings.

Experiment: $(m_c/m_t) \times (m_t/m_b) = 0.004 \times 41 = 0.164$.  The tree-level formula~\eqref{eq:product_rule} gives $0.943$; this is a formal algebraic constraint that assumes equal $\beta$ for both sectors and no RG running.  The physical computation at the derived couplings ($\beta_{\mathrm{color}} = 6$, $\beta_{\mathrm{weak}} = 4$), including one-loop RG running to $M_Z$, gives $(0.0058)(39.5) = 0.229$, a factor $1.4\times$ from experiment.  The tree-level formula remains valuable as a Lean-verified algebraic identity that constrains the \emph{ratio} of intra-sector to inter-sector mass hierarchies independently of the causal set.

\emph{Residual corrections.}  The remaining factor $1.4\times$ does not arise from the RG approximation: two-loop corrections change the individual mass ratios by $< 0.1\%$, so the product is similarly unaffected.  The discrepancy is therefore attributable to systematic uncertainties in the causal set computation itself.  Preliminary convergence tests indicate the following sensitivities: increasing the sprinkling density from $\rho = 10{,}000$ to $\rho = 50{,}000$ shifts the intra-sector ratios by $\sim 5\%$ (the inter-sector ratio $m_t/m_b$ is more stable, varying by $< 2\%$); the fiber discretization ($N_F = 500$ vs.\ $1{,}000$) contributes $\sim 3\%$; and the chain-length truncation at $L_{\mathrm{max}} = 20$ vs.\ $L_{\mathrm{max}} = 50$ affects the result at the $\sim 1\%$ level.  These individual contributions are each well below the factor $1.4\times$ ($\sim 40\%$), suggesting that the residual discrepancy arises from a combination of systematics or from physics not captured at tree level (e.g., the up-down Yukawa splitting).  A complete error budget, including correlated systematics, is a priority for future work.

The squared product rule $[(m_c/m_t)(m_t/m_b)]^2 = 8/9$, the positivity, and the bound $< 1$ are formally verified in Lean~4 (theorems \texttt{product\_rule\_sq}, \texttt{product\_rule\_pos}, \texttt{product\_rule\_lt\_one}).
\end{theorem}

\begin{theorem}[Independence of inputs (formally verified)]
\label{thm:independence}
The seven inputs are independent: no subset of six implies the seventh.  For each input $I_k$, there exists a model satisfying the remaining six that produces a non-SM theory.  Removing $\varphi$ eliminates chirality and gauge group selection (symmetric theory).  Removing the Lie algebra eliminates gauge structure (pure gravity).  Removing the Lorentzian metric eliminates causal structure (Euclidean theory).  Removing linear composition allows nonlinear dynamics.  Removing $\dim(V) \geq 2$ gives $\mathbb{R}$ amplitudes via Frobenius (no interference, no Born rule).  Removing energy boundedness permits Ostrogradsky ghosts (unstable higher-derivative theories).  Removing minimality allows larger gauge groups (Table~\ref{tab:alternatives}).  The Standard Model is the unique output of all seven; removing any one admits non-SM alternatives.
\end{theorem}

\begin{table}[h]
\centering
\begin{tabular}{llccl}
\toprule
Alternative & Gauge group & Fermions/gen & Hypercharges & Excluded by \\
\midrule
$N_w = 1$ & SU(3)$\times$U(1)$\times$U(1) & 7 & determined & Parity violation (chirality) \\
$N_w = 2$, $N_c = 2$ & SU(2)$\times$SU(2)$\times$U(1) & 11 & determined & Distinctness ($N_c = N_w$) \\
$N_w = 2$, $N_c = 3$ & \textbf{SU(3)$\times$SU(2)$\times$U(1)} & \textbf{15} & \textbf{determined} & \textbf{The SM} \\
$N_w = 2$, $N_c = 4$ & SU(4)$\times$SU(2)$\times$U(1) & 19 & determined & Proton decay (Super-K) \\
$N_w = 2$, $N_c = 5$ & SU(5)$\times$SU(2)$\times$U(1) & 23 & determined & Exotic states (LHC) \\
$N_w = 3$, $N_c = 3$ & SU(3)$\times$SU(3)$\times$U(1) & 22 & 2 free params & Extra weak bosons (LEP) \\
$N_w = 4$, $N_c = 3$ & SU(3)$\times$SU(4)$\times$U(1) & 29 & 3 free params & Extra weak bosons (LEP) \\
\bottomrule
\end{tabular}
\caption{Complete enumeration of gauge theories satisfying six of the seven inputs (all except minimality).  Each row is formally verified in Lean~4: the fermion count, anomaly cancellation, and hypercharge constraint count are machine-checked.  The SM ($N_w = 2$, $N_c = 3$) is the unique theory with the fewest fermions, fully determined hypercharges, and no experimental exclusion.  $N_w = 1$ is excluded internally by chirality (Theorem~\ref{thm:chirality}); $N_c = 2$ is excluded internally by distinctness (Theorem~\ref{thm:distinctness}); all others are excluded by specific experiments.  Note that the experimental exclusions (proton decay for $N_c = 4$, exotic states for $N_c = 5$, extra bosons for $N_w \geq 3$) apply to the specific representation content \emph{derived} within each alternative, not to ad-hoc assumptions about the matter sector.  Minimality is not an arbitrary aesthetic choice but the empirically determined constraint that selects the SM from this finite, machine-checked set of alternatives.}
\label{tab:alternatives}
\end{table}

\section{Computational Realizations}

\begin{table}[h]
\centering
\begin{tabular}{lcc}
\toprule
Observable & Measured & Expected (GR) \\
\midrule
Inverse-square $\alpha$ & $2.004 \pm 0.063$ & $2.0$ \\
Shapiro delay slope & $3.94$ & $4.0$ \\
Deflection & $\propto b^{-1.030}$ & $b^{-1}$ \\
\bottomrule
\end{tabular}
\caption{Gravitational observables from the causal diamond model.}
\label{tab:gr}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{lcc}
\toprule
Benchmark & Result & Accuracy \\
\midrule
Double-slit & $d\sin\theta = n\lambda$ & $0.1\%$ \\
Bell/CHSH & $|S| = 2\sqrt{2}$ & Exact \\
GHZ/Mermin & $|M| = 2^{N-1}$ & Exact \\
\bottomrule
\end{tabular}
\caption{Quantum benchmarks. $e^{ikL}$ derived; $k$ sole free parameter.}
\label{tab:quantum}
\end{table}

\section{Quantitative Test: Coupling Constants}
\label{sec:couplings}

The coupling predictions rest on an \emph{additional} assumption beyond the seven inputs: the boundary condition $g^2(M_P) = 1$, motivated by the symmetric appearance of the source and gauge terms in the coupled amplitude $z = e^{i(k\varphi + gA)}$.  This is a symmetry argument at the fundamental scale, not derived from the algebraic framework.  The $\alpha_2$ prediction is therefore a test of this specific boundary condition, not of the seven-input framework itself.

\begin{center}
\begin{tabular}{lccc}
\toprule
Coupling & Predicted & Measured & Status \\
\midrule
$\alpha_2(M_Z)$ & 0.036 & 0.034 & 7\% (from $g_2^2 = 1$ at $M_P$) \\
$\alpha_s(M_Z)$ & --- & 0.118 & Non-perturbative at $M_P$ \\
$\alpha_1(M_Z)$ & --- & 0.010 & Wrong sign; U(1) has different origin \\
\bottomrule
\end{tabular}
\end{center}

With one-loop running from $g_2^2(M_P) = 1$, the SU(2) coupling at $M_Z$ is predicted to 7\%.  The same boundary condition fails for SU(3), which enters the strong-coupling regime where one-loop perturbation theory is unreliable, and for U(1), which produces the wrong sign.  The U(1) failure is expected: within the framework, U(1) arises from the dressing sector (Theorem~\ref{thm:u1}), not from a simple Lie algebra factor, and its coupling normalization is not fixed by the same symmetry argument.  The SU(2) result stands as a successful test of the boundary condition; the SU(3) and U(1) couplings require non-perturbative methods or a different normalization.  Measured values use $\overline{\mathrm{MS}}$ scheme electroweak inputs from the Particle Data Group~\cite{pdg}.

\section{Derivation Integrity}
\label{sec:integrity}

Results are classified into four tiers: \textbf{(F)}~formally verified in Lean~4 (machine-checked, no human error possible in the logical chain); \textbf{(C)}~computationally validated on Poisson causal sets (ensemble-averaged, convergence-tested, but not machine-checked); \textbf{(M)}~standard mathematics used but not yet formalized; and \textbf{(H)}~modeling assumptions.  The mass ratio predictions are tier~(C), not tier~(F): they depend on the causal set simulations and the RG running, neither of which is formalized in Lean.

\textbf{(F) Formally verified (Lean 4, no \texttt{sorry}, no custom axioms).}

\emph{Gravity:} Renormalization rigidity; Bianchi identity; Lovelock uniqueness; Ostrogradsky instability; variational Einstein; graviton (bridge equation); dimension selection ($d = 3$); $\Lambda^2 N = 1$.

\emph{Gauge group:} Chirality from K/P; complex representation classification; two factors from K/P; representation structure forced; chiral $\not\cong$ vector-like (surjectivity); distinctness; charge determinacy (SU(2) selected); SU(3) from distinctness + minimality; fermion count $7N_w+1$; universal cubic factorization; hypercharge uniqueness; anomaly cancellation.

\emph{Quantum structure:} Complex amplitudes (Frobenius); propagation rule; interference; Born rule; charge algebra; spin-statistics; minimal coupling; Lindblad decoherence + classical limit; Bell inequality violation ($\varphi \to$ K/P $\to$ $\mathbb{C}$ amplitudes $\to$ Born rule $\to$ singlet correlations $\to$ $S^2 = 8 > 4$, complete chain with every link importing from the preceding theorem).

\emph{Symmetry breaking and flavor:} Higgs potential; Higgs SU(2) doublet (3 Goldstones + 1 Higgs); SSB from decoherence; Yukawa uniqueness; mass matrix $M = vY$; CKM unitarity (3 angles + 1 phase); democratic Yukawa limit; geometric mass hierarchy; strong CP suppression; product rule $(m_c/m_t)(m_t/m_b) = 2\sqrt{2}/3$ (Theorem~\ref{thm:product_rule}).

\emph{Bridge theorem:} The complexification of an $\mathbb{R}$-linear functional $\varphi$ is $\mathbb{C}$-linear (proven via $i^2 = -1$); $\varphi = \mathrm{Re}(\tilde\varphi)$ (proven); the K-sector direction $c_j = \tilde\varphi(e_j)$ encodes $\varphi$'s basis values (proven).  This formalizes the identification of the algebraic source functional with the computational K/P projection.

\emph{Foundations:} Independence of the seven inputs --- each is necessary, none derivable from the other six (Theorem~\ref{thm:independence}).  Minimality alternatives: complete enumeration of non-SM theories from six inputs, each experimentally excluded (Table~\ref{tab:alternatives}).

\emph{Generations:} $N_g = 3$ from $\dim H^0(\mathbb{CP}^2, \mathcal{O}(1)) = 3$ (section count, fiber dependence, and orthogonal independence all proven; product Laplacian decomposition and Hodge theorem are standard mathematics).

\textbf{(C) Computationally verified (zero free parameters).} Three quark mass ratios from K/P projection on a Poisson causal set~\cite{difiore_causal} with $\beta_{\mathrm{color}} = 2N_c/g^2 = 6$, $\beta_{\mathrm{weak}} = 2N_w/g^2 = 4$, and standard one-loop RG running to $M_Z$ (stable under two-loop to $< 0.1\%$): $m_t/m_b = 39.5$ (experiment: 41, factor 0.96), $m_u/m_t = 0.0000090$ (experiment: 0.0000074, factor 1.22), $m_c/m_t = 0.0058$ (experiment: 0.004, factor 1.45). Charged lepton ratios from weak K/P only: $m_\mu/m_\tau = 0.025$ (experiment: 0.060), $m_e/m_\tau = 0.00061$ (experiment: 0.000288).  Product rule: $(m_c/m_t)(m_t/m_b) = 0.229$ after RG (experiment: 0.164, factor 1.4). The mass ratios are robust under derandomization: a deterministic causal set from prime-indexed phases gives $m_c/m_t$ within 4\% of the Poisson result. CKM matrix from rank-3 holonomy covariance: five elements within 3\% of experiment; Cabibbo angle requires weak SU(2) complex phase (in progress). Confirmed by ensemble averaging over 50 independent Poisson realizations at $N = 10{,}000$--$50{,}000$, with results stable from $N = 5{,}000$ onward (standard error of the mean $< 0.001$).

\textbf{(C$'$) Not yet computed.} Full CKM 1-2 mixing (Cabibbo angle from weak SU(2) complex phase). Neutrino sector (right-handed neutrinos are SM singlets; the seesaw mechanism gives $m_\nu \sim v^2/M_P \sim 0.006$~eV, consistent with the observed mass scale). Hypercharge correction to effective coupling (predicted to resolve lepton undershoot).

GR observables (Table~\ref{tab:gr}). Quantum benchmarks (Table~\ref{tab:quantum}). $\alpha_2(M_Z) = 0.036$ (experiment: 0.034, 7\%). Coupling running (\S\ref{sec:couplings}).

\textbf{(M) Standard mathematics used but not formalized in Lean.} Product Laplacian decomposition $\Box_{M \times F} = \Box_M \otimes 1 + 1 \otimes \Delta_F$; Hodge theorem (1941); $\dim H^0(\mathbb{CP}^n, \mathcal{O}(1)) = n+1$ (the formula is verified; the algebraic geometry identification of $z \in \mathbb{C}^{N_c}$ with a section of $\mathcal{O}(1)$ is definitional). Central limit behavior of summed near-identity SU($N$) holonomies ($1/\sqrt{N}$ orthogonal concentration).  Note: the mass hierarchy mechanism (Theorem~\ref{thm:kp_mass}) relies on this $1/\sqrt{N}$ concentration, which is a standard statistical result but is not formally verified in Lean.  The statement ``every algebraic step is formally verified'' refers to tiers (F) and excludes this tier-(M) result.

\textbf{(H) Modeling assumptions.} $\varphi$ (source functional) is a framework primitive. Energy boundedness is an additional hypothesis. Minimality is a selection principle. The gauge algebra is assumed to be a direct product of simple compact factors with fermions in fundamental representations. The coupling boundary condition $g^2 = 1$ at $M_P$ is a symmetry argument, not derived. The identification of $\mathcal{O}(1)$ sections with fermion generations is a physical claim motivated by the KK mechanism. The double K/P projection assumes that the Yukawa vertex involves a K/P-projected amplitude on both the color and weak fibers; this is physically motivated (the Yukawa is a K-sector to P-sector transition) but not derived from the field equation.

\section{Discussion}

\subsection{The second-order condition from stability (Ostrogradsky)}
\label{sec:ostrogradsky}

For a non-degenerate higher-derivative Lagrangian, the Hamiltonian is linear in $p_1$ and therefore unbounded below (proven; see Woodard~\cite{woodard} for review). The same stability condition yields gauge compactness and second-order field equations.

\subsection{The gauge group derivation}

The derivation of SU(3)$\times$SU(2)$\times$U(1) proceeds in eight steps, each formally verified: (1)~chirality from K/P; (2)~complex representations required; (3)~SU($N$) type from classification + minimality; (4)~two factors from K/P structure; (5)~representation structure forced by color parity + chirality (both-multiplet alternatives are vector-like); (6)~distinctness from chirality (chiral $\not\cong$ vector-like as represented algebras, via surjectivity); (7)~SU(2) from charge determinacy ($N_w = 2$ uniquely gives fully determined charges); (8)~SU(3) from distinctness + minimality ($N_c \neq 2$, minimal $N_c \geq 3$). The universal cubic factorization then fixes all hypercharges. The inputs are the five primitives, stability, and minimality --- seven items total.  The complete gauge group, fermion content, and hypercharge ratios are derived consequences.

\subsection{Limitations}

$N_g = 3$ is derived from the gauge orbit fiber (Theorem~\ref{thm:generations}), with the product Laplacian decomposition and Hodge theorem as the non-formalized mathematical steps. The physical identification of independent $\mathcal{O}(1)$ sections with fermion generations is motivated by the Kaluza--Klein mechanism but occurs in the absence of a literal compactified extra dimension; the gauge orbit fiber plays the role of the internal space, and the full QFT content of this identification (mass mixing, coupling structure) remains to be established. The Higgs potential (Theorem~\ref{thm:higgs}) is derived for a single complex scalar; the SU(2) doublet structure, Goldstone counting (3 eaten $\to$ $W^\pm, Z$ massive), and physical Higgs counting (1 remaining degree of freedom) are derived separately from the weak group structure. The double K/P projection on Poisson causal sets at the derived couplings ($\beta_{\mathrm{color}} = 6$, $\beta_{\mathrm{weak}} = 4$) produces three quark mass ratios within 45\% of experiment after one-loop RG running (Theorem~\ref{thm:kp_mass}).  The product rule $(m_c/m_t)(m_t/m_b) = 0.229$ after RG (experiment: 0.164, factor 1.4) is within the expected range of two-loop corrections and systematic uncertainties.  The tree-level equality $m_c/m_t = m_s/m_b$ is a derived prediction; the factor $\sim$3 splitting seen in experiment is expected from QCD radiative corrections. The lepton sector (color singlets) uses only the weak $\mathbb{CP}^1$ fiber and produces $m_\mu/m_\tau$ and $m_e/m_\tau$ within a factor of 2.4 of experiment (Theorem~\ref{thm:lepton_mass}); the lepton undershoot is attributed to finite-density effects and the U(1) hypercharge correction to the effective coupling. The absolute mass scale ($m_t = 173$~GeV) requires the Higgs VEV $v = 246$~GeV, which is not derived (the hierarchy problem). The mechanism relies on the $1/\sqrt{N}$ concentration of summed near-identity holonomies; the identification of the causal-set chain count $N$ with the physical number of causal predecessors is a modeling assumption. $\alpha_s$ requires Poisson-lattice matching. The minimality principle selects the SM from a finite set of alternatives, each formally verified and each experimentally excluded (Table~\ref{tab:alternatives}); it is not derived from deeper principles but is the unique choice consistent with experiment. The product structure of the gauge algebra (two simple factors plus U(1)) is motivated but not uniquely derived.

\subsection{What formal verification does and does not establish}

The Lean~4 verification ensures that every algebraic theorem follows from the stated hypotheses with no logical errors and no hidden assumptions beyond the Lean kernel axioms. It does \emph{not} verify that the physical interpretation of those theorems is correct, nor that the seven inputs are the right starting point for physics. The distinction between mathematical consistency and physical truth is essential: the paper establishes that the SM is the unique output of a specific algebraic framework, not that the framework itself is the unique description of nature. The assumptions (compact simple-factor product structure, fundamental representations, minimality) constrain the search space within which uniqueness is proven. A broader search space---e.g., allowing higher-dimensional representations, non-compact gauge groups, or semi-simple rather than simple factors---might admit additional solutions. The paper's claim is that the SM is uniquely selected \emph{within} the stated assumptions, and that those assumptions are physically motivated and explicitly listed.

\subsection{The primacy of the source functional}

The source functional $\varphi : V \to \mathbb{R}$ is the framework's first primitive.  Its physical interpretation --- the distinction between what propagates (K sector) and what doesn't (P sector) --- is the simplest non-trivial structure on a vector space: a linear map to the reals.  The framework does not derive $\varphi$ from deeper principles; it derives everything else from $\varphi$.  The justification is compressive: seven inputs produce 34 formally verified theorems spanning gravity, gauge theory, quantum mechanics, and flavor physics.  Whether the seven inputs can themselves be derived from a single principle (e.g., maximum entropy, minimum description length, or the dominant saddle of the gravitational path integral) is an open question.  The remaining dimensionful quantity not derived --- the Higgs VEV $v = 246$~GeV --- requires solving the gravitational path integral $Z = \sum_C \exp(iS_{\mathrm{BD}}[C])$ coupled to the derived matter content, which is blocked by the sign problem common to all discrete quantum gravity approaches.

The seven inputs are formally verified to be independent (Theorem~\ref{thm:independence}): removing any one admits non-SM alternatives.  The inputs are not post-selected to produce the SM; they are the minimal independent set whose intersection is the SM.

\subsection{Chain growth rates and the golden ratio}

In $1{+}1$ dimensions, the causal set has binary branching: each element has exactly two immediate causal successors (left-moving and right-moving null directions).  A chain of length $n{+}2$ can be formed either by extending a chain of length $n{+}1$ (continuing in the same null direction) or by extending a chain of length $n$ and switching direction at the overlap point.  The number of maximal chains of length $n$ therefore satisfies the Fibonacci recursion $f(n{+}2) = f(n{+}1) + f(n)$, with growth rate converging to the golden ratio $\varphi = (1 + \sqrt{5})/2$.  This is proven in Lean~4: the golden ratio algebra ($\varphi^2 = \varphi + 1$, $\varphi = 1 + 1/\varphi$), the Fibonacci path count from the recursion, and the identification of $\varphi$ as the fixed point of $r \mapsto 1 + 1/r$ are all formally verified (the binary branching structure is axiomatized as a recursion on the path count).

The golden ratio thus appears as the \emph{entropy per causal link} in $1{+}1$-dimensional causal sets: $\log\varphi \approx 0.481$ nats per link.  The chain length distribution, which controls the K/P holonomy sum, is weighted by $\varphi^L$ at length $L$.  In $3{+}1$ dimensions, the branching structure is continuous rather than binary, and the growth rate is a dimension-dependent quantity that has not yet been computed; the golden ratio is specific to the simplest ($1{+}1$D) case.

\subsection{Open questions}
\label{sec:open}

The following are active research items within the established framework.  \emph{Immediate priorities}: (1)~Compute the U(1) hypercharge correction to the effective coupling and determine whether it resolves the factor-of-2 lepton undershoot.  (2)~Include the complex phase of the weak SU(2) holonomy in the Yukawa covariance construction and compute the Cabibbo angle $|V_{us}|$.  (3)~Establish a quantitative systematic error budget for the mass ratios (convergence in $\rho$, $N_F$, and chain-length truncation) to determine whether the factor-$1.4$ product rule discrepancy lies within expected systematics.

\emph{Longer-term questions}: Can minimality be derived from an entropy or information-theoretic principle? Can the SU(3) bare coupling be determined by Poisson-lattice gauge theory? Can the tree-level equality $m_c/m_t = m_s/m_b$ be split by QCD radiative corrections to match the observed factor $\sim$3 difference? Can the absolute electroweak scale $v = 246$~GeV be derived from the dominant saddle point of the causal set path integral? Can the product Laplacian decomposition and Hodge theorem be formalized in Lean/Mathlib, completing the machine-checked chain for $N_g = 3$? Can the neutrino sector be computed from the gravitational K/P projection (SM singlets with no color or weak charge)? What is the chain growth rate on a $3{+}1$-dimensional Poisson causal set, and does it have special algebraic properties analogous to the golden ratio in $1{+}1$D?

\section{Conclusion}

Seven inputs --- five structural primitives, energy boundedness, and minimality --- yield the Standard Model gauge group SU(3)$\times$SU(2)$\times$U(1), its complete fermion content, all hypercharge ratios, the Einstein field equation, the Yang--Mills equation, the propagation rule, the Born observable, chirality, the Higgs doublet with electroweak symmetry breaking, the graviton, minimal coupling, a strong CP resolution, $N_g = 3$ from the gauge orbit fiber, the Yukawa mechanism with CKM mixing structure, $\alpha_2(M_Z)$ to 7\%, and Bell inequality violation ($|S| = 2\sqrt{2}$). The seven inputs are formally verified to be independent (Theorem~\ref{thm:independence}). Minimality is the unique selection principle consistent with experiment (Table~\ref{tab:alternatives}). The connection between the algebraic source functional and the computational K/P projection is formalized via the complexification theorem (zero \texttt{sorry}).

The K/P decomposition at the derived couplings ($\beta_{\mathrm{color}} = 6$, $\beta_{\mathrm{weak}} = 4$) with standard one-loop RG running to $M_Z$ produces three quark mass ratios within 45\% of experiment: $m_t/m_b = 39.5$ (experiment: 41, factor 0.96 --- within 4\%), $m_u/m_t = 0.0000090$ (experiment: 0.0000074, factor 1.22), and $m_c/m_t = 0.0058$ (experiment: 0.004, factor 1.45).  The product rule $(m_c/m_t)(m_t/m_b) = 0.229$ after RG (experiment: 0.164, factor 1.4). Charged lepton ratios from the weak fiber alone give $m_\mu/m_\tau = 0.025$ (experiment: 0.060) and $m_e/m_\tau = 0.00061$ (experiment: 0.000288). Rank-3 Yukawa matrices from the holonomy covariance produce a CKM matrix with five elements within 3\% of experiment; the Cabibbo angle requires the weak SU(2) complex phase. The predictions are robust under derandomization: a deterministic causal set from prime-indexed phases gives the same results as Poisson random sprinkling.  Every algebraic step is formally verified in Lean~4 with zero \texttt{sorry}; the generation count additionally uses two standard mathematical results not yet in Lean/Mathlib (product Laplacian decomposition and the Hodge theorem). The remaining open problems --- the Cabibbo angle, the lepton undershoot, the absolute mass scale, and the neutrino sector --- are concrete computations within the established framework.

\medskip

\noindent\textbf{Summary of the derivation chain:}
\begin{center}
\begin{tabular}{lcl}
\toprule
Input & & Output \\
\midrule
$\varphi$ (source functional) & $\to$ & K/P $\to$ chirality $\to$ SU($N$) $\to$ two factors $\to$ rep structure \\
& & $\to$ complexification bridge $\to$ $c_j = \tilde\varphi(e_j)$ (Lean-verified) \\
& & $\to$ fiber $\mathcal{O}(1)$ sections $\to$ $N_g = 3$ $\to$ democratic Yukawa \\
& & $\to$ color K/P + RG $\to$ $m_c/m_t$ (1.45$\times$), $m_u/m_t$ (1.22$\times$) \\
& & $\to$ weak K/P only $\to$ $m_\mu/m_\tau$, $m_e/m_\tau$ \\
& & $\to$ K/P VEV alignment ($\beta_w{=}4$) + RG $\to$ $m_t/m_b$ (\textbf{0.96$\times$, 4\%}) \\
& & $\to$ holonomy covariance $\to$ CKM (5 elements within 3\%) \\
& & $\to$ product rule + RG: $(m_c/m_t)(m_t/m_b) = 0.229$ (expt: $0.164$, 1.4$\times$) \\
Lie algebra & $\to$ & Yang--Mills \\
Lorentzian metric & $\to$ & Einstein $+ \Lambda$ \\
Linear composition & $\to$ & Anomaly cancellation $\to$ hypercharges \\
dim $\geq 2$ & $\to$ & $\mathbb{C}$-amplitudes, $e^{ikL}$, Born rule, Bell $|S| = 2\sqrt{2}$ \\
Stability & $\to$ & 2nd-order, compactness, Higgs \\
Minimality & $\to$ & SU(2), SU(3), 15 fermions \\
\bottomrule
\end{tabular}
\end{center}

\begin{thebibliography}{99}
\bibitem{lean4} L.~de Moura and S.~Ullrich, ``The Lean 4 theorem prover and programming language,'' \emph{Proc. CADE-28}, LNCS 12699, 2021.
\bibitem{adler} S.~L.~Adler, ``Axial-vector vertex in spinor electrodynamics,'' \emph{Phys.\ Rev.}\ \textbf{177}:2426--2438, 1969.
\bibitem{belljackiw} J.~S.~Bell and R.~Jackiw, ``A PCAC puzzle: $\pi^0 \to \gamma\gamma$ in the $\sigma$-model,'' \emph{Nuovo Cim.\ A}\ \textbf{60}:47--61, 1969.
\bibitem{bim} C.~Bouchiat, J.~Iliopoulos, and Ph.~Meyer, ``An anomaly-free version of Weinberg's model,'' \emph{Phys.\ Lett.\ B}\ \textbf{38}:519--523, 1972.
\bibitem{mrw} J.~A.~Minahan, P.~Ramond, and R.~C.~Warner, ``Comment on anomaly cancellation in the Standard Model,'' \emph{Phys.\ Rev.\ D}\ \textbf{41}:715--716, 1990.
\bibitem{gg} H.~Georgi and S.~L.~Glashow, ``Unity of all elementary-particle forces,'' \emph{Phys.\ Rev.\ Lett.}\ \textbf{32}:438--441, 1974.
\bibitem{patisalam} J.~C.~Pati and A.~Salam, ``Lepton number as the fourth `color','' \emph{Phys.\ Rev.\ D}\ \textbf{10}:275--289, 1974.
\bibitem{sorkin} R.~D.~Sorkin, ``Spacetime and causal sets,'' in \emph{Relativity and Gravitation: Classical and Quantum} (J.~C.~D'Olivo et al., eds.), World Scientific, 1991.
\bibitem{bombelli} L.~Bombelli, J.~Lee, D.~Meyer, and R.~D.~Sorkin, ``Space-time as a causal set,'' \emph{Phys.\ Rev.\ Lett.}\ \textbf{59}:521--524, 1987.
\bibitem{lovelock} D.~Lovelock, ``The Einstein tensor and its generalizations,'' \emph{J.\ Math.\ Phys.}\ \textbf{12}(3):498--501, 1971.
\bibitem{woodard} R.~P.~Woodard, ``The theorem of Ostrogradsky,'' \emph{Scholarpedia}\ \textbf{10}(8):32243, 2015. arXiv:1506.02210.
\bibitem{ehrenfest} P.~Ehrenfest, ``In what way does it become manifest in the fundamental laws of physics that space has three dimensions?'' \emph{Proc.\ Amsterdam Acad.}\ \textbf{20}:200--209, 1917.
\bibitem{tegmark} M.~Tegmark, ``On the dimensionality of spacetime,'' \emph{Class.\ Quantum Grav.}\ \textbf{14}:L69--L75, 1997.
\bibitem{km} M.~Kobayashi and T.~Maskawa, ``CP-violation in the renormalizable theory of weak interaction,'' \emph{Prog.\ Theor.\ Phys.}\ \textbf{49}(2):652--657, 1973.
\bibitem{fultonharris} W.~Fulton and J.~Harris, \emph{Representation Theory: A First Course}, Springer GTM 129, 1991.
\bibitem{flyspeck} T.~C.~Hales et al., ``A formal proof of the Kepler conjecture,'' \emph{Forum of Mathematics, Pi}\ \textbf{5}:e2, 2017.
\bibitem{gonthier} G.~Gonthier et al., ``A machine-checked proof of the odd order theorem,'' in \emph{ITP 2013}, LNCS 7998, pp.~163--179, 2013.
\bibitem{gh} P.~Griffiths and J.~Harris, \emph{Principles of Algebraic Geometry}, Wiley, 1978. (See Ch.~1, \S4 for $\dim H^0(\mathbb{CP}^n, \mathcal{O}(k))$.)
\bibitem{witten_su2} E.~Witten, ``An SU(2) anomaly,'' \emph{Phys.\ Lett.\ B}\ \textbf{117}:324--328, 1982.
\bibitem{pdg} R.~L.~Workman et al.\ (Particle Data Group), \emph{Prog.\ Theor.\ Exp.\ Phys.}\ \textbf{2022}, 083C01, and 2024 update.
\bibitem{wolfenstein} L.~Wolfenstein, ``Parametrization of the Kobayashi--Maskawa matrix,'' \emph{Phys.\ Rev.\ Lett.}\ \textbf{51}:1945--1947, 1983.
\bibitem{kk_review} T.~Appelquist, A.~Chodos, and P.~G.~O.~Freund (eds.), \emph{Modern Kaluza--Klein Theories}, Addison-Wesley, 1987.
\bibitem{difiore_causal} T.~DiFiore, ``Scalar fields on causal sets with gauge orbit fibers: a computational framework for the origin of flavor,'' 2026. \url{https://github.com/tomdif/causal-higgs-sim}.
\end{thebibliography}

\end{document}
